### 2023-04-29 papers/beyond-may.tex
reorganize and edit draft

\documentclass[
 prl,
 twocolumn,
 amsmath,
 amssymb,
 aps,
]{revtex4-2}

\usepackage{colortbl}
\usepackage{graphicx} 
\usepackage{booktabs}
\usepackage{caption}
\usepackage{notes2bib}
\newcommand{\red}{\textcolor{red}}
\usepackage[justification=raggedright]{caption}

\begin{document}

\title{Beyond May: Complexity-stability relationships \\
in disordered dynamical systems}

\author{Onofrio Mazzarisi}
\affiliation{Department of Ecology and Evolutionary Biology, University of Kansas, Haworth Hall, 
1200 Sunnyside Avenue
Lawrence, KS 66045, USA}

\author{Matteo Smerlak}
\affiliation{Laboratoire Biophysique et Evolution, ESPCI ParisTech, 10 rue Vauquelin, 75005 Paris, France}
\affiliation{Capital Fund Management, 23 Rue de l'Université, 75007 Paris, France}

\date{\today}% It is always \today, today,
             %  but any date may be explicitly specified

\begin{abstract}

\end{abstract}
\maketitle

\section*{Introduction}

\section*{Results}

\paragraph*{\bf Model.}
\input{model.tex}

\paragraph*{\bf Homogeneous interactions.}
\input{homogeneous.tex}

\paragraph*{\bf Random interactions: general reults.}
\input{random-general.tex}
\paragraph*{\bf Random interactions: power laws case.}
\input{random-power.tex}

\section{Conclusion}
\input{conclusion.tex}

\begin{acknowledgments}
We gratefully acknowledge multiple discussions with Ian Hatton, Matthieu Barbier and Ada Altieri on ecological power laws and the diversity-stability debate. 
\end{acknowledgments}

\clearpage

\bibliography{general-stability}
\bibliographystyle{unsrt}

\end{document}