### 2023-04-29 papers/conclusion.tex

The relationship between complexity and stability in high-dimensional dynamical systems has been a longstanding puzzle.

We reconsidered the relationship between diversity and stability in complex systems, a classic problem since May's 1973's puzzling prediction that "complexity begets instability".

Considering dynamical models more general than the classical Lotka-Volterra model, we found that the relation can in fact go both ways:  depending on the relative convexity of self- and cross-interactions,