### 2023-05-05 papers/beyond-may.tex
update draft

\documentclass[
 prl,
 twocolumn,
 amsmath,
 amssymb,
 aps,
]{revtex4-2}

\usepackage{colortbl}
\usepackage{graphicx} 
\usepackage{booktabs}
\usepackage{caption}
\usepackage{notes2bib}
\newcommand{\red}{\textcolor{red}}
\usepackage[justification=raggedright]{caption}

\captionsetup{justification=Justified}

\begin{document}

\title{Beyond May: Complexity-stability relationships \\
in disordered dynamical systems}

\author{Onofrio Mazzarisi}
\affiliation{Department of Ecology and Evolutionary Biology, University of Kansas,
Higuchi Hall, 2101 Constant Avenue, Lawrence, KS 66047, USA}

\author{Matteo Smerlak}
\affiliation{Laboratoire Biophysique et Evolution, ESPCI ParisTech, 10 rue Vauquelin, 75005 Paris, France}
\affiliation{Capital Fund Management, 23 Rue de l'Université, 75007 Paris, France}

\date{\today}

\begin{abstract}
\input{abstract.tex}
\end{abstract}
\maketitle

%Introduction
\paragraph*{}
\input{intro.tex}

%Model
\paragraph*{}
\input{model.tex}

%Homogeneous interactions
\paragraph*{}
\input{homogeneous.tex}

%Random interactions: stability condition
\paragraph*{}
\input{stability.tex}

%Random interactions: equilibrium distribution
\paragraph*{}
\input{equilibrium.tex}

%Conclusion
\paragraph*{}
\input{conclusion.tex}

\medskip

\begin{acknowledgments}
We thank Ian Hatton, Matthieu Barbier and Ada Altieri for productive discussions on ecological power laws and the diversity-stability debate and Vadim Karatayev and Daniel Reuman for feedbacks on the manuscript.
O.M. was partly supported by NSF BIO OCE grant 2023473.
\end{acknowledgments}

\medskip

\bibliography{beyond-may}
\bibliographystyle{unsrt}

\end{document}