### 2023-05-07 papers/beyond-may.tex
review and edit changes

\documentclass[
 prl,
 twocolumn,
 amsmath,
 amssymb,
 aps,
]{revtex4-2}

\usepackage{colortbl}
\usepackage{graphicx} 
\usepackage{booktabs}
\usepackage{caption}
\usepackage{notes2bib}
\newcommand{\red}{\textcolor{red}}
\usepackage[justification=raggedright]{caption}

\captionsetup{justification=Justified}

\begin{document}

\title{Beyond May: Complexity-stability relationships \\
in disordered dynamical systems}

\author{Onofrio Mazzarisi}
\affiliation{Department of Ecology and Evolutionary Biology, University of Kansas,
Higuchi Hall, 2101 Constant Avenue, Lawrence, KS 66047, USA}

\author{Matteo Smerlak}
\affiliation{Laboratoire Biophysique et Evolution, ESPCI ParisTech, 10 rue Vauquelin, 75005 Paris, France}
\affiliation{Capital Fund Management, 23 Rue de l'Université, 75007 Paris, France}

\date{\today}

\begin{abstract}
\input{abstract.tex}
\end{abstract}
\maketitle

%Introduction
\paragraph*{}
\input{intro.tex}

%Model
\paragraph*{}
\input{model.tex}

%Homogeneous interactions
\paragraph*{}
\input{homogeneous.tex}

%Random interactions: stability condition
\paragraph*{}
\input{stability.tex}

%Random interactions: equilibrium distribution
\paragraph*{}
\input{equilibrium.tex}

%Conclusion
\paragraph*{}
\input{conclusion.tex}

\medskip

\begin{acknowledgments}
We thank Ian Hatton, Matthieu Barbier and Ada Altieri for productive discussions on ecological power laws and the diversity-stability debate, and Vadim Karatayev and Daniel Reuman for feedback on the manuscript.
Funding for this work was provided by the Alexander von Humboldt Foundation in the framework of the Sofja Kovalevskaja Award endowed by the German Federal Ministry of Education and Research.
O.M. was partly supported by NSF BIO OCE grant 2023473.
\end{acknowledgments}

\medskip

\bibliography{beyond-may}
\bibliographystyle{unsrt}

\end{document}