### 2023-05-07 papers/conclusion.tex

\emph{Discussion.---}The relationship between complexity and stability in high-dimensional dynamical systems has been a longstanding puzzle, in ecology and in other fields. 
In this letter, we have showed that the condition $\sigma\sqrt{N}< \mu_s - \mu$ (popularized by May's slogan ``complexity begets instability") does not provide a complete picture of the relationship between complexity and stability. 
In particular, we have seen that generalized Lotka-Volterra model, often cited in support of May's prediction, corresponds to a special cancellation in the more general stability condition $N\langle (\sigma/\psi)^2\rangle < 1$.
In models where self-interactions grow faster than cross-interactions, the opposite behavior is observed: stability becomes \emph{more} likely with increasing diversity $N$.

Future work can look at the case where two-body interactions are not separable, or include higher order interaction to generalize the results in Ref. \cite{Gibbs2022} based on the GLV model.